---
title: "Zeyd Khalil HW8, October 15, 2020"
output: pdf_document
---

```{r, warning=FALSE}
library(tidyverse)
```  

# Exercise 1 - The following built-in datasets are not tidy. For each one, describe why it is not tidy and write out what the first five entries would look like once it is in a tidy format.

> a. relig_income
> b. billboard
> c. us_rent_income

```{r}
head(relig_income, n = 5)
```  

> The reason why this data is not tidy is because the column headings are values and not variable names. In order to tidy this data, I would use count() in order to distribute the income intervals in each row, and count the number of cases in those intervals. The result would consist of 3 columns; religion, income interval, and count. That would significantly increase the number of rows, but decrease the number of columns. 

```{r}
head(billboard, n = 5)
```  

> The reason why this data is not tidy is the same as in the previous example, because most of the columns in this dataset are numerical values instead of variable names. In order to tidy this data, I would use count() in order to distribute the week number in each row, and count the number of cases in those intervals. The result would consist of 4 columns; artist, track, date entered, week number, and count. That would significantly increase the number of rows, but decrease the number of columns. 

```{r}
head(us_rent_income, n = 5)
```  

> The reason why this data is not tidy is because each observational unit is spread across multiple rows. In order to make this data tidy, what I would do is use mutate() in order to create a new column for rent. Once that's done, the first five entries of the data will consist of 5 columns and only one row per state; the columns will be GEOID, NAME, Income, Rent, and moe. 

# Exercise 2 - Try on your own to do the same thing to table4b. 

```{r}
tidy4b <- table4b %>% pivot_longer(cols = c(`1999`, `2000`), names_to = "year", values_to = "cases")

tidy4b
```  

# Exercise 3 - Tidy built-in dataset relig_income  

```{r}
tidy_relig_income <- relig_income %>% pivot_longer(cols = c(`<$10k`, `$10-20k`, `$20-30k`, `$30-40k`, `$40-50k`, `$50-75k`, `$75-100k`, `$100-150k`, `>150k`, `Don't know/refused`), names_to = "religion income", values_to = "count")

tidy_relig_income
```  

# Exercise 4 - 

```{r}
monkeymem <- read_csv("https://raw.githubusercontent.com/JaneWall/data_STAT412612/master/monkeymem.csv")

tidy_monkeymem <- monkeymem %>% pivot_longer(cols = c(Week2, Week4, Week8, Week12, Week16), names_to = "Week", values_to = "percent")

tidy_monkeymem
```  

# Exercise 5 - Tidy the fish_encounters dataset of fish spotting by monitoring stations. Make the NA into 0 using the option values_fill = list(seen = 0)

```{r}
tidy_fishencounters <- fish_encounters %>% pivot_wider(names_from = station, values_from = seen, values_fill = list(seen = 0))

tidy_fishencounters
```   

# Exercise 6 - Spread the flowers1 data frame. Hint: use read_csv2() to read in the dataset.

```{r}
flowers1 <- read_csv2("https://raw.githubusercontent.com/JaneWall/data_STAT412612/master/flowers1.csv")

spread_flowers1 <- flowers1 %>% pivot_wider(names_from = Variable, values_from = Value)

spread_flowers1
```  

# Exercise 7 - Tidy the dataset flowers2.csv by turning the one column into 3 separate columns.

```{r}
flowers2 <- read_csv("https://raw.githubusercontent.com/JaneWall/data_STAT412612/master/flowers2.csv")

sep_flowers2 <- flowers2 %>% separate(`Flowers/Intensity;Time`, into = c("Flowers", "Intensity", "Time"), sep = "/") %>% separate(Intensity, into = c("Intensity", "Time"))  

sep_flowers2
```  

# Exercise 8 - Re-unite the data frame you separated from the flowers2 exercise. Use a comma for the separator.  

```{r}
unite_flowers2 <- sep_flowers2 %>% unite(`Flowers,Intensity`, Flowers, Intensity, sep = ",")

unite_flowers2
```  

# Exercise 9 - In the following dataset, turn the implicit missing values to explicit.  

```{r}
output <- tibble(
      treatment   = c("a", "b", "a", "c", "b"),
      gender   = factor(c("M", "F", "F", "M", "M"), levels = c("M", "F", "O")),
      return = c(1.5, 0.75,  0.5,  1.8,  NA)
    )
    output

output %>% complete(treatment, gender)
```

# Exercise 10 - Use pivot_longer() to put the days all in one column. Then, rearrange the data.

```{r}
weather <- read_csv("https://raw.githubusercontent.com/JaneWall/data_STAT412612/master/weather.csv")

weather_tidy <- weather %>% pivot_longer(cols = starts_with("d"), names_to = "day", names_pattern = "d(.*)", values_to = "value") %>% na.omit()

weather_tidy

weather4 <- weather_tidy %>% pivot_wider(names_from = element, values_from = value)

summary(weather4)
```  

# Exercise 11 - Tidy the billboard dataset (built-in)

* First gather up all the week entries into a row for each week for each song (where there is an entry)

* Then, convert the week variable to a number and figure out the date corresponding to each week on the chart

* Sort the data by artist, track and week.

```{r}
billboard %>% pivot_longer(cols = starts_with("wk"), names_to = "week", names_pattern = "wk(.*)", values_to = "rank") %>% na.omit() %>% select(artist, track, week, rank, date.entered)

billboard$week <- as.integer(billboard$week)
```